---
title: "lab5"
author: 
  - Xuan Wen
  - Roy Khoo
date: "2023-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)
library(lubridate)
```

```{r}
# Load accident.csv file
accident <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/accident.csv", stringsAsFactors = FALSE)

states <- map_data("state")
states %>% ggplot(aes(x = long, y = lat)) + geom_path(aes(group = group)) + 
  geom_point(aes(x = LONGITUD, y = LATITUDE), size = 0.05, color = 'red', data = accident %>% filter(YEAR == 2016, LONGITUD < 0, LONGITUD > -130))
```

# Part one: Accident data

1. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

```{r}
accident_day = accident %>% mutate(DAY_WEEK = factor(DAY_WEEK)) %>%
  group_by(DAY_WEEK) %>% summarise(count = n())

accident_day %>% ggplot(aes(x = DAY_WEEK)) + geom_bar(aes(weight = count))
# Monday and the weekends has generally more accident counts than other days.
```

2. What time of the day do accidents happen (use variable HOUR)?

```{r}
accident %>% filter(HOUR <= 24) %>% group_by(HOUR) %>% summarise(count = n()) %>% ggplot(aes(x = HOUR)) +
  geom_bar(aes(weight = count))

# 6.00 pm has the highest rates of accident, possibly due to it being the time most people drive back from work
```

3. What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
count(accident %>% filter(DRUNK_DR > 0), NULL)
# 8720
```


# Part two: Connecting data

1. Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.

2. Join accident and driver table (work out which variable(s) to use)

3. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
